---
title: "PISA Predictors: Data Analysis"
author: "Edoardo Ruvolo, Sami Louahlia"
format: html
   # wordcount-html:
   #  toc: true
   #  fig-align: center
   #  cap-location: margin
execute:
  message: false
  warning: false
echo: false
---


```{r}
here::i_am("Data_Project_M1_Edoardo_Sami.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(nlstools)
library(gtsummary)
library(corrr)
library(maps)
```
## Research Project : Exploratory analysis of potential predictors of PISA Scores

[Link to the Github Repository](https://github.com/SamiLhl/Data_Project_M1_Edoardo_Sami) 
This is the same Repository as for the Data Treatment, we now proceed with the Dataset we produced in a new quarto document.

## Part 1 : Research Question 

The national level of education is an important element to account for in public policy. In economic terms, it represents how well the nation develops its human capital, offering greater economic outcomes to its inhabitants.  Universal access to education may also contribute to improving the equality of opportunity within a country’s population. Moreover, beyond resource production and distribution, people may value education as a good in itself. 
The OECD’s PISA initiative is a remarkable opportunity for cross-country comparisons in terms of educational attainments at the lower secondary level, when students are aged around 15. It is also a major opportunity to use cross-country differences in other metrics to determine potential predictors for greater national PISA scores.

  Using PISA datasets that reported national scores from 2009 to 2018, along with variables gathered across multiple international organizations, this project searches for potential variables, able to explain cross-country differences in performances at the PISA test. The variables used are different in nature. Some feature the current state of a nation’s economy, others observe their demographic condition, while others tackle the way education is provided in the different observed countries to a deeper extent. As such, we hope to acquire a wide array of possible predictors for higher or lower PISA scores.

  The completion of this research project may provide valuable insights. It may offer explanations as to why some countries perform better than others, orienting the agency of respective national governments. It may also leave room for some form of determinism, as some variables, such as geographic, demographic, or economic variables, may impact students’ performances while being more difficult or less desirable for governments to act upon in the short term.  
Finally, this research project provides potential criticism as to what the PISA score truly reflects. Some variables have been selected to test potential flaws as to how the OECD can gather information about a nation's educational attainment, which could overlook important elements such as who can be evaluated by the OECD when devising the PISA scores and do they accurately represent the general student population of the studied country.

## Part 2 : Data sets descriptions (size, production condition, main variables)*



## Part 3 : Data analysis

```{r}
Dataset <- read.csv(here("Final_Data"))
Dataset <- Dataset |> select(-Jurisdiction.x) |> rename(Country = Jurisdiction.y)
```
```{r}

# Load world map data
world_map <- map_data("world")

filtered_MAP <- Dataset #|>
  #filter(Country %in% world_map$region)|> filter(Year==2018)

# Merge your data with the map data
filtered_MAP <- filtered_MAP |>
    mutate(Country = ifelse(Country == "United States", "USA", Country))|>
    mutate(Country = ifelse(Country == "Viet Nam", "Vietnam", Country))|>
    mutate(Country = ifelse(Country == "United Kingdom", "UK", Country))
PISA_map <- left_join(world_map, filtered_MAP, by = c("region" = "Country"))


# Create the ggplot world map with color-coded values
ggplot(PISA_map, aes(x = long, y = lat, group= group, fill = Science_Score)) +
  geom_polygon(color = "white") +
  scale_fill_gradient(low = "yellow", high = "red", na.value = "grey",
    guide = guide_colorbar(title = "Science Score")) +
  labs(title = "Map of PISA Science Scores in 2018") +
  theme_void()
```
Before starting analysis it is important to acknowledge some important limitations of the PISA data. As this map shows, important parts of the world do not participate to the PISA test. While Europe, North and South America as well as Oceania are almost fully represented in the PISA dataset, only few countries of the Arab world and South East Asia participated, and no Sub-Saharan nor south Asian countries were represented in the Dataset. As such, most of the developing world is not taken into account. Thus, the number of independent observations is limited, justifying our usage of up to four observations per countries, among those that participated to the PISA test. Additionally, the studied variables may lack the variability needed to fully observe variables effects on the PISA score. Insights that developing countries could have potentially offered.


```{r}
Dataset |> group_by(Year) |> summarise(Math_Score = mean(Math_Score), Science_Score=mean(Science_Score), Reading_Score = mean(Reading_Score)) |>
ggplot (aes(x = Year)) +
  geom_line(aes(y = Math_Score, color = "Math Score")) +
  geom_line(aes(y = Science_Score, color = "Science Score")) +
  geom_line(aes(y = Reading_Score, color = "Reading Score")) +
  xlab("Year") +
  ylab("Score") +
  ggtitle("Scores Growth Over Years") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("Math Score" = "blue", "Science Score" = "red", "Reading Score" = "green"))
```
In order to have a clearer understanding of the data, we started by plotting the mean of the score over the years, and we discovered that they present a similar pattern. There is a clear drop after 2012, that caused scores to drop by 20 points circa. Since we found the three scores having a similar behaviors, we decided to conduct our analysis only using one of them for simplicity in the readability of the results. 

In our case the science score seems the most interesting to take into consideration. 
```{r Function tabling stuff}

Model_to_Table <- function(Model){

# Get the summary table
lm_summary <- summary(Model)

# Extract coefficients table
coefficients_table <- lm_summary$coefficients
}
```
## Financial Condition and PISA score

Rich countries may generally be expected to provide better education services, as they have greater means for doing so, and need a highly educated population to perform tasks with important added value. Economies with an ambition for reaching high income status may also strive towards educating its population to become as productive as in advanced economies. Thus, observing these indicators in their ability to predict PISA scores is of great relevance. Moreover, these characteristics, if powerful in predicting PISA scores, might be used as control variables to observe and isolate the impact of other socio-economic phenomena, later on , in our analysis. Therefore, we start our analysis by looking at these variables. 

### GNI per Capita

```{r}
ggplot(Dataset, aes(x= GNI.per.capita, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue")+
xlab("GNI per capita") +
ylab("Science_Score") +
ggtitle("Scatterplot of PISA Science score given GNI per capita") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
It appears that GNI per capita is powerfully correlated with PISA scores. However, it seems that it ceases to be, past a certain level of income. At first sight, this could perhaps be represented by a form of a log function, given we overlook the very high income outliers, which represent small populations, such as Qatar, Singapore or Luxembourg. For simplicity, we stick to a linear relationship in modelling GNI per capita predicting PISA Science score.

```{r GNI and PISA}
mod_GDP_sci <- lm(Science_Score ~ GNI.per.capita, data=Dataset)
lmGNI <- Model_to_Table(mod_GDP_sci)
Table_GNI <- knitr::kable(lmGNI)
```


`r Table_GNI` GNI per capita is indeed, significantly associated with higher Science Scores (p<0.001).
For every additional dollar, the Science score increases by 0.001, so that an additional 1000 dollar is associated with one additional point on the PISA score. NO RSQUARE AS OF NOW


### GDP growth and PISA Scores 
Booming economies may have hopes to become an advanced economy in the span of a generation or two, in a similar fashion to the Asian tigers did in past. In a similar fashion that South Korea consolidated after its economic miracle, and joined the ranks of advanced economies, through notably human capital investment (Koen et al.,2021), countries experiencing high economic growth may put greater efforts in their education services, and thus, have greater PISA scores. In dealing with this variable we rid ourselves of the 2009 observations taking place at the same time as the Global Financial Crisis.

```{r}
NO2009 <- Dataset |> filter(Year != 2009)
ggplot(NO2009, aes(x= GDP.rate, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue")+
xlab("GDP growth rate") +
ylab("Science_Score") +
ggtitle("Scatterplot of PISA Science score given GDP growth rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

As can be seen, graphical representation of Science Score being only predicted by growth rate does not show much. By doing so, we confound high income countries, which have lower growth rate as they do not experience growth from convergence with some lower income countries that did not successfully entered economic convergence yet, also having low growth rate. Or as we said with the previous analysis, GNI per capita is associated with higher PISA scores. Thus, it is necessary to control for GNI per capita before assessing the relevance of the growth rate of the economy in predicting PISA scores.

```{r gy controlling for GNI per cap}

mod_gy_sci_GDPc <- lm(Science_Score ~ GDP.rate + GNI.per.capita , data=Dataset)
lmgy_GNIc <- Model_to_Table(mod_gy_sci_GDPc)
Table_gy_GNI_all <- knitr::kable(lmgy_GNIc)

```
`r Table_gy_GNI_all`


Another strategy might be to try and isolate low and middle income countries. As of 2023, a nation is considered part of the High Income group once the GDP per capita reaches US$13845.

```{r gy controlling for only low_mid inc GNI per cap}
NO2009Rich <- NO2009 |> filter(GNI.per.capita <= 13845)
mod_gy_sci_GDPc <- lm(Science_Score ~ GDP.rate + GNI.per.capita , data=NO2009Rich)
lmgy_GNIc <- Model_to_Table(mod_gy_sci_GDPc)
Table_gy_GNI <- knitr::kable(lmgy_GNIc)

```
`r Table_gy_GNI`Once controlling for GDP per capita, it appears that the growth rate of the economy is not significantly correlated with PISA scores. GDP per capita is not significant either in this model. However, it is important to acknowledge that only 29 observations had GNI per capita below US$13845, making any statistical investigation hardly feasible.

## Demographic indicators and PISA scores

### Fertility Rates and PISA Scores


The literature, generally supports a negative relationship between fertility rates and educational achievements. One possible explanation takes the form of a 'quality over quantity' reasoning. When the population rapidly expands in size, the quantity of workers, able to finance children's education, relative to the quantity of children, in need of education provision, diminishes (Hu, 2023; Sittisom, 2020). This is also notably channelled through women enjoying greater education than their parents, having less children. For instance, Osili and Long (2007) found that increase female education by one year reduces early fertility by 0.26 in Nigeria. This implies that it might also be education that reduces fertility rate (Brand and Davis, 2012). 
As such, we expect countries with high fertility rates to have lower PISA scores. 

```{r Fert_Sci}
ggplot(Dataset, aes(x= Fertility.Rate, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue")+
xlab("Fertility Rate") +
ylab("Science_Score") +
ggtitle("Scatterplot of PISA Science score given Fertility Rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There seem to be some form of negative linear relationship between fertility rate and PISA scores. 
```{r first attempt with Fertility}
mod_fer_sci <- lm(Science_Score ~ Fertility.Rate, data = Dataset)
lmfer <- Model_to_Table(mod_fer_sci)
Table_Fer_Sci <- knitr::kable(lmfer)
```
`r Table_Fer_Sci`

The correlation is significant (p<.01). The best linear model provide sizeable coefficient. For an increase in fertility rate of 1, the expected national PISA score decreases 55.1 in science.

It may be important to account for the fact that low fertility rates at a national level are generally associated with higher living standards and higher income. As such, it might be necessary to control for some metrics, such as gdp per capita to ensure that fertility rates truly play a role in providing better education.

```{r Controlling for GDP per capita}
mod_fer_sci_GDPc <- lm(Science_Score ~ Fertility.Rate + GNI.per.capita, data=Dataset)
lmfer_GDPc <- Model_to_Table(mod_fer_sci_GDPc)
Table_Fer_GDPc <- knitr::kable(lmfer_GDPc)
```
`r Table_Fer_GDPc`

After controlling for GNI per capita, fertility is still significantly and negatively associated with PISA scores(p<.001). An increase of 1 in fertility rate is associated with a 43.85 points decrease in Science score.

It might be challenging, however, to disentangle how fertility affects education and education affect fertility, out of PISA data, ???as we lack the ability to determine which of the two variables had precedence in their change???. 

### Population Density 

Population density could potentially affect academic performance through multiple channels. For instance, low population density could lead to longer commuting time to school. However, high density areas may have multiple schools and reinforce social segregation between schools in turn potentially decreasing the means available to children of lower strata people. Borland and Howsen (1999) investigate this dynamic in Kentucky and notice that students from highly urban and highly rural areas perform similarly, and worse than students enjoying more moderate population density.

```{r Dens_Sci}
ggplot(Dataset, aes(x= Poulation.density, y=Science_Score))+
geom_point()+
xlab("Population Density") +
ylab("Science_Score") +
ggtitle("Scatterplot of PISA Science score given Population Density") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

There does not seem to be any form of linear correlation between population density and PISA scores. There seem to be little to support any non-linear relationship either.

### Entrance Age

There are important debate on the impact of the entrance age of pupils into school and their academic performances. The Child Care Research Network (2007) argued that even though, kids entering school at an older age did perform better than their younger counterparts, these advantages were small in magnitude and decreased in importance over time. 
```{r}
Entrance_PISA <- Dataset |> select(Science_Score, Entrance_Mandatory, GNI.per.capita)|> filter(!is.na(Entrance_Mandatory))
```


```{r Entr_Sci}
ggplot(Entrance_PISA, aes(x= as.factor(Entrance_Mandatory), y=Science_Score))+
geom_boxplot()+
xlab("Mandatory Entrance Age") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science score given national Entrance Age into school") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There seem to be a non-linear relationship between Mandatory Entrance Age and PISA scores. Which reflects quite well the tradeoffs often mentioned in the literature, according to which, children may need time to mature before entering school, even though waiting for too long might not be profitable as children would be ready for being schooled and were not schooled, meaning they might reach the age/year for taking the PISA with less competencies acquired at school than other pupils who started school earlier.()
For simplicity however, we only realize a linear model of Entrance Age and PISA scores, as the maximum mandatory entrance age observe was 7, which does not seem, at least visually, to negatively impact students too much, relative to 6 years mandatory entrance age. Were the data to contain countries with higher age of mandatory entrance, using a linear model could have been more problematic.  

```{r}
Entrance_PISA$Entr_M <- as.numeric(Entrance_PISA$Entrance_Mandatory)
mod_Entr <- lm(Science_Score ~ Entr_M + GNI.per.capita, data=Entrance_PISA)
lmEntr <- Model_to_Table(mod_Entr)
Table_Entrance <- knitr::kable(lmEntr)

```
`r Table_Entrance`

Even while controlling for GDP per capita, mandatory entrance age is here significantly associated with PISA scores (p<.001). Making entry to school mandatory one year later is associated with a 15.54 points increase in PISA score.
It is important to point out that the data is a reflection of the policies carried out by governments about the age at which children are required to attend school. These do not necessarily reflect the actual entrance age in studied countries, only the approach of the government on the matter.


## Mean Yrs of Schooling MIGHT HAVE TO NONO OUT OF EXISTENCE FOR NOT BEING RELEVANT AT ALL (THE LADS ARE ALL TAKING THE TEST IN LATE MIDDLE SCHOOL SO THAT SHOWS NOTHING BEYONG MEAN YRS BEING HIGHER FOR RICH COUNTRIES)

At first sight, the national mean years of schooling may seem as a curious variable to observe. As students are all tested at the same level namely end of middle school, normally aged 15. However, the fact that students are expected to complete many more years in schooling could lead programs to be less dense.  
```{r Mean_Schl_Math}
ggplot(Dataset, aes(x= Mean_Yrs_Schooling, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue")+
xlab("Mean Years of Schooling") +
ylab("Math Score") +
ggtitle("Scatterplot of PISA Reading score given Mean Years of schooling ") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
As expected, there seem to be a positive correlation between mean years of schooling. We now wonder whether this positive tendency is due only to other factors, while mean years of schooling actually affects negatively performances at the PISA test.

```{r}
mod_MYrs <- lm(Science_Score ~ Mean_Yrs_Schooling + GNI.per.capita + Fertility.Rate, data=Dataset)
lmMYrs <- Model_to_Table(mod_MYrs)
Table_Myrs <- knitr::kable(lmMYrs)
```
`r Table_Myrs`

It does not seem as Mean Years of Schooling can explain more than what is already explained by GNI per capita. When pooling them into a model, GNI per capita becomes insignificant. We favour keeping GNI per capita as it a more logical indicator when observing PISA scores of children who only reached end of middle school at the time of tests.


## Academic Year Length

Some of the skills acquired at school may be lost as a result of extended period without practising those skills. As such, part of the literature argues that long summer holidays may negatively affect student performances as mentioned by Finnie et al. (2019) Thus, we try looking for differences between countries where the academic year is relatively shorter and those where it is relatively longer
```{r}
YrLength_PISA <- Dataset |> select( Science_Score, End_Month, Start_Month)
```

```{r}
YrLength_PISA$Year_Length <- (YrLength_PISA$End_Month - YrLength_PISA$Start_Month + 12) %% 12
YrLength_PISA$Year_Length <- ifelse(YrLength_PISA$Year_Length < 4, NA, YrLength_PISA$Year_Length)
YrLength_PISA$Year_Length_binary <- ifelse(YrLength_PISA$Year_Length <=9 , "Short", "Long")
```
We omit some outliers with very small school years.

```{r Yr_Length_Sci}
YrLength_PISA <- YrLength_PISA |> filter(!is.na(Year_Length))
#ggplot(YrLength_PISA, aes(x= as.factor(Year_Length), y=Science_Score))+
ggplot(YrLength_PISA, aes(x= Year_Length_binary, y=Science_Score))+
  geom_boxplot()+
xlab("Year_Length") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science score given the Year Length") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
At first sight, there does not seem to be a notable difference between short and long years.

```{r}
chi_Yr <- chisq.test(YrLength_PISA$Year_Length_binary, YrLength_PISA$Science_Score)
#Sum_Chi_Yr <- tbl_summary(chi_Yr)
#Table_Chi_Yr <- knitr::kable(Sum_Chi_Yr)
#`r Table_Chi_Yr`
```
A Chi-Squarred test confirms that year length is not significant in predicting the value of the PISA score. Having a longer or a shorter break in between two grades does not appear to affect students' performances.


```{r}
Dataset %>% select(Science_Score, Survival_Rate_Primary, Gov_Expenditure, Enrolment_Rate, Share_Private) %>% correlate()
```


### Survival rate 

```{r Mean_Surv_Math}
ggplot(Dataset, aes(x=Survival_Rate_Primary, y=Science_Score))+
geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
xlab("Survival rate") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science score given the mortality rate ") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r first attempt  sur}
mod_sur_sci <- lm(Science_Score ~ Survival_Rate_Primary, data = Dataset)
lmsur <- Model_to_Table(mod_sur_sci)
Table_Sur_Sci <- knitr::kable(lmsur)
```
`r Table_Sur_Sci`

```{r}
mod_surrs <- lm(Science_Score ~ Survival_Rate_Primary + GNI.per.capita, data=Dataset)
lmsurrs <- Model_to_Table(mod_surrs)
Table_Sur_Sci1 <- knitr::kable(lmsurrs)
```
`r Table_Sur_Sci1`

a comment is to be added about the results 

### Government Expenditure

```{r}
ggplot(Dataset, aes(x=Gov_Expenditure, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
xlab("Governamenent Expenditure") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science score given gvernament expenditure") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Science score is independent from the governament expenditure. The average governament expenditure is around 1 and is related to a math score of 490.
By increasing the expenditure there is no benefit to the math score.

```{r first attempt gov }
mod_gov_sci <- lm(Science_Score ~ Gov_Expenditure, data = Dataset)
lmgov <- Model_to_Table(mod_gov_sci)
Table_Gov_Sci <- knitr::kable(lmgov)
```
`r Table_Gov_Sci`

a comment is to be added about the results (not significant because p>0.1)

### Enrolment rate
```{r}
ggplot(Dataset, aes(x=Enrolment_Rate, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
xlab("Enrolemnt Rate") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science Score given Enrolment Rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r first attempt ER }
mod_er_sci <- lm(Science_Score ~ Enrolment_Rate, data = Dataset)
lmer <- Model_to_Table(mod_er_sci)
Table_Er_Sci <- knitr::kable(lmer)
```
`r Table_Er_Sci`

```{r}
mod_ersr <- lm(Science_Score ~ Enrolment_Rate + GNI.per.capita, data=Dataset)
lmer1 <- Model_to_Table(mod_ersr)
Table_Er_Sci1 <- knitr::kable(lmer1)
```
`r Table_Er_Sci1`

A comment is to be added

### Share of private expenditure
```{r}
ggplot(Dataset, aes(x=Share_Private, y=Science_Score))+
geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
xlab("Share Private") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science Score given Share Private") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r first attempt sp }
mod_sp_sci <- lm(Science_Score ~ Share_Private, data = Dataset)
lmsp <- Model_to_Table(mod_sp_sci)
Table_Sp_Sci <- knitr::kable(lmsp)
```
`r Table_Sp_Sci`

significance at p<0.1
```{r}
mod_sp <- lm(Science_Score ~ Share_Private + GNI.per.capita, data=Dataset)
lmsp1 <- Model_to_Table(mod_sp)
Table_Sp_Sci1 <- knitr::kable(lmsp1)
```
`r Table_Sp_Sci1`

A comment is to be added

### Trial on the regression 
```{r}
lm1 <- lm(Science_Score ~ GNI.per.capita + Fertility.Rate + Entrance_Mandatory + Survival_Rate_Primary + Enrolment_Rate + Share_Private, data=Dataset)
summary(lm1)
```
drop share private

```{r}
lm2 <- lm(Science_Score ~ GNI.per.capita + Fertility.Rate + Entrance_Mandatory + Survival_Rate_Primary + Enrolment_Rate, data=Dataset)
summary(lm2)
```


Add a comment for each of them, and we need to properly apply the function

## Annex + Sources


### Sources 

Child Care Research Network, N. E. (2007). Age of Entry to Kindergarten and Children’s Academic Achievement and Socioemotional Development. Early Education and Development, 18(2), 337. https://doi.org/10.1080/10409280701283460

Borland, M. V., & Howsen, R. M. (1999). A Note on Student Academic Performance: In Rural versus Urban Areas. The American Journal of Economics and Sociology, 58(3), 537–546. http://www.jstor.org/stable/3487782

Hu, H. The Consequences of Fertility Decline on Educational Attainment in China. Popul Res Policy Rev 42, 94 (2023). https://doi-org.proxy.bu.dauphine.fr/10.1007/s11113-023-09834-7

Sittisom,W. (2020). Barriers in way towards quality education: poverty, fertility rate and
increasing population, Journal of Security and Sustainability Issues, 9(January), 215-227. http://doi.org/10.9770/jssi.2020.9.J(16)

Brand, J. E., & Davis, D. (2011). The Impact of College Education on Fertility: Evidence for Heterogeneous Effects. Demography, 48(3), 863. https://doi.org/10.1007/s13524-011-0034-3

Osili, U. and Long, B. T. (2007). Does female schooling reduce fertility? evidence from nigeria.. https://doi.org/10.3386/w13070

Finnie, R. K. C., Peng, Y., Hahn, R. A., Johnson, R. L., Fielding, J. E., Truman, B. I., Muntaner, C., Fullilove, M. T., Zhang, X., & Community Preventive Services Task Force (2019). Examining the Effectiveness of Year-Round School Calendars on Improving Educational Attainment Outcomes Within the Context of Advancement of Health Equity: A Community Guide Systematic Review. Journal of public health management and practice : JPHMP, 25(6), 590–594. https://doi.org/10.1097/PHH.0000000000000860

Koen, V. André, C. Beom, J. Purwin A. Byungjun K.,(2021) Better Policies to Prolong the Miracle on the Han River, OECD, 82-95.   