---
title: "PISA Predictors: Data Analysis"
author: "Edoardo Ruvolo, Sami Louahlia"
format: html
execute:
  message: false
  warning: false
echo: false
---


```{r}
here::i_am("Data_Project_M1_Edoardo_Sami.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(nlstools)
```
## Research Project : Exploratory analysis of potential predictors of PISA Scores

[Link to the Github Repository](https://github.com/SamiLhl/Data_Project_M1_Edoardo_Sami) 
This is the same Repository as for the Data Treatment, we now proceed with the Dataset we produced in a new quarto document.

## Part 1 : Research Question 





## Part 2 : Data sets descriptions (size, production condition, main variables)*



## Part 3 : Data analysis

```{r}
Dataset <- read.csv(here("Final_Data"))
```
## Financial Condition and PISA score
Rich countries may generally be expected to provide better education services as they have greater means for doing so and need a highly educated population to perform tasks with important added value. As a result. Economies with an ambition for reaching high income status may also strive towards educating its population to become as productive as in advanced economies. Thus, observing these indicators in their ability to predict PISA scores is of great relevance. Moreover, these characteristics, if powerful in predicting PISA score, might be used as control variable to observe other socio-economic phenomena. Therefore, we start our analysis with these variables. 

### GNI per Capita and Fertility Rate

```{r}
ggplot(Dataset, aes(x= GNI.per.capita, y=Math_Score))+
geom_point()+
xlab("GNI per capita") +
ylab("Math_Score") +
ggtitle("Scatterplot of PISA Math score given GNI per capita") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
It appears that GNI per capita is powerfully correlated with PISA scores. However, it seems that it ceases to be, past a certain level of income. At first sight, this could perhaps be represented by a form of a log function, given we overlook the very high income outliers, which generally represent small population such as Qatar, Singapore or Luxembourg.

```{r GNI and PISA}
mod_GDP_mat <- lm(Math_Score ~ GNI.per.capita, data=Dataset)
```


### GDP growth and PISA Scores 
Booming economies may have hopes to become an advanced economy in the span of a generation or two, in a similar fashion the Asian tigers did in past. As such countries experiencing high economic growth may put greater efforts in their education services and have greater PISA scores.

```{r}
ggplot(Dataset, aes(x= GDP.rate, y=Math_Score))+
geom_point()+
xlab("GDP growth rate") +
ylab("Math_Score") +
ggtitle("Scatterplot of PISA Math score given GDP growth rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
As can be seen, graphical representation of Math Score being only predicted by growth rate does not show much. By doing so, we confound high income countries, which have lower growth rate as they do not experience growth from convergence with lower income countries that also have lower growth rate. It is necessary to control for GNI per capita before assessing the relevance of the growth rate of the economy in predicting PISA scores.

```{r gy controlling for GNI per cap}
mod_gy_mat_GDPc <- lm(Math_Score ~ GDP.rate + GNI.per.capita , data=Dataset)
summary(mod_gy_mat_GDPc)
```
Once controlling for GDP per capita, it appears that the growth rate of the economy is not significantly correlated with PISA scores.


### Fertility Rates and PISA Scores

The underlying assumption behind suspecting this relation stems from a 'quality over quantity' reasoning. When the population rapidly expands in size, the quantity of workers, able to finance children's education, relative to the quantity of children, in need of education provision, diminishes. As such, we expect countries with high fertility rates to have lower PISA scores. 

```{r Fert_Math}
ggplot(Dataset, aes(x= Fertility.Rate, y=Math_Score))+
geom_point()+
xlab("Fertility Rate") +
ylab("Math_Score") +
ggtitle("Scatterplot of PISA Math score given Fertility Rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r Fert_Sci}
ggplot(Dataset, aes(x= Fertility.Rate, y=Science_Score))+
geom_point()+
xlab("Fertility Rate") +
ylab("Science_Score") +
ggtitle("Scatterplot of PISA Science score given Fertility Rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r Fert_Read}
ggplot(Dataset, aes(x= Fertility.Rate, y=Reading_Score))+
geom_point()+
xlab("Fertility Rate") +
ylab("Reading_Score") +
ggtitle("Scatterplot of PISA Reading score given Fertility Rate") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There seem to be some form of negative linear relationship between fertility rate and PISA scores. The Reading Score seem to correlate less strongly with fertility rates. 

```{r first attempt with Fertility}
cor_fer_mat <- cor.test(Dataset$Math_Score,Dataset$Fertility.Rate)
mod_fer_mat <- lm(Math_Score ~ Fertility.Rate, data = Dataset)

cor_fer_sci <- cor.test(Dataset$Science_Score,Dataset$Fertility.Rate)
mod_fer_sci <- lm(Science_Score ~ Fertility.Rate, data = Dataset)

cor_fer_read <- cor.test(Dataset$Reading_Score,Dataset$Fertility.Rate)
mod_fer_read <- lm(Reading_Score ~ Fertility.Rate, data = Dataset)

##table_fert <- summarise(co 

```

All the correlations are significant (p<.01). The best linear models provide sizeable coefficients. For an increase in fertility rate of 1, the expected national PISA score decreases by 62.4 in math, 55.1 in science and 52.2 in reading.
It may be important to account for the fact that low fertility rates at a national level are generally associated with higher living standards and higher income. As such, it might be necessary to control for some metrics, such as gdp per capita to ensure that fertility rates truly play a role in providing better education.

```{r Controlling for GDP per capita}
mod_fer_mat_GDPc <- lm(Math_Score ~ Fertility.Rate + GNI.per.capita, data=Dataset)
mod_fer_sci_GDPc <- lm(Science_Score ~ Fertility.Rate + GNI.per.capita, data=Dataset)
mod_fer_red_GDPc <- lm(Reading_Score ~ Fertility.Rate + GNI.per.capita, data=Dataset)
summary(mod_fer_mat_GDPc)
summary(mod_fer_sci_GDPc)
summary(mod_fer_red_GDPc)
```
After controlling for GNI per capita, fertility are still, significantly, negatively associated with PISA scores. An increase of 1 in fertility rate is associated with a 48 points decrease in PISA Math scores, a 43.85 points decrease in Science and a 40.08 decrease in Reading Score.

### Population Density 


```{r Dens_Math}
ggplot(Dataset, aes(x= Poulation.density, y=Math_Score))+
geom_point()+
xlab("Population Density") +
ylab("Math_Score") +
ggtitle("Scatterplot of PISA Math score given Population Density") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r Dens_Sci}
ggplot(Dataset, aes(x= Poulation.density, y=Science_Score))+
geom_point()+
xlab("Population Density") +
ylab("Science_Score") +
ggtitle("Scatterplot of PISA Science score given Population Density") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r Dens_Read}
ggplot(Dataset, aes(x= Poulation.density, y=Reading_Score))+
geom_point()+
xlab("Fertility Rate") +
ylab("Reading_Score") +
ggtitle("Scatterplot of PISA Reading score given Population Density") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
There does not seem to be any form of linear correlation between population density and PISA scores. 

### Entrance Age
There are important debate on the impact of the entrance age of pupils into school and their academic performances. 
```{r}
Entrance_PISA <- Dataset |> select(Math_Score, Reading_Score, Science_Score, Entrance_Mandatory)|> filter(!is.na(Entrance_Mandatory))
```


```{r Entr_Math}

ggplot(Entrance_PISA, aes(x= as.factor(Entrance_Mandatory), y=Math_Score))+
geom_boxplot()+
xlab("Mandatory Entrance Age") +
ylab("Math Score") +
ggtitle("Scatterplot of PISA Math score given national Entrance Age into school") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r Entr_Sci}
ggplot(Entrance_PISA, aes(x= as.factor(Entrance_Mandatory), y=Science_Score))+
geom_boxplot()+
xlab("Mandatory Entrance Age") +
ylab("Science Score") +
ggtitle("Scatterplot of PISA Science score given national Entrance Age into school") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r Entr_Read}
ggplot(Entrance_PISA, aes(x= as.factor(Entrance_Mandatory), y=Reading_Score))+
geom_boxplot()+
xlab("Mandatory Entrance Age") +
ylab("Reading Score") +
ggtitle("Scatterplot of PISA Reading score given national Entrance Age into school") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There seem to be a non-linear relationship between Mandatory Entrance Age and PISA scores. Which reflects quite well the tradeoffs often mentioned in the literature, according to which, children may need time to mature before entering school, even though waiting for too long might not be profitable as children would be ready to for being schooled and are not, meaning they might reach the age/year for taking the PISA with less competencies acquired during school than other pupils who started school earlier.()
We thus try modeling a non linear relationship between Mandatory Entrance Age and PISA scores

```{r}
##mod_age_mat <- nls(Math_Score ~ log(Entrance_Mandatory), data =Entrance_PISA, start = ##(Entrance_Mandatory = 3))
```


## Mean Yrs of Schooling MIGHT HAVE TO NONO OUT OF EXISTENCE FOR NOT BEING RELEVANT AT ALL (THE LADS ARE ALL TAKING THE TEST IN LATE MIDDLE SCHOOL SO THAT SHOWS NOTHING BEYONG MEAN YRS BEING HIGHER FOR RICH COUNTRIES)
```{r Mean_Schl_Math}
ggplot(Dataset, aes(x= Mean_Yrs_Schooling, y=Math_Score))+
geom_point()+
xlab("Mean Years of Schooling") +
ylab("Math Score") +
ggtitle("Scatterplot of PISA Reading score given Mean Years of schooling ") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Start and End Month

This variable may be used in multiple fashions. One point of interest might be the length of the school year. 
(ADD ARTICLES SAYING STUFF ABOUT THAT)
```{r}
YrLength_PISA <- Dataset |> select(Math_Score, Reading_Score, Science_Score, End_Month, Start_Month)
```

```{r}
YrLength_PISA$Year_Length <- (YrLength_PISA$End_Month - YrLength_PISA$Start_Month + 12) %% 12
YrLength_PISA$Year_Length <- ifelse(YrLength_PISA$Year_Length < 4, NA, YrLength_PISA$Year_Length)
```
Some values seem hardly probable and are very rare, thus we omit them for further analysis (EXPAND ON THAT)

```{r Yr_Length_Math}
YrLength_PISA <- YrLength_PISA |> filter(!is.na(Year_Length))
ggplot(YrLength_PISA, aes(x= as.factor(Year_Length), y=Math_Score))+
geom_boxplot()+
xlab("Year_Length") +
ylab("Math Score") +
ggtitle("Scatterplot of PISA Math score given the Year Length") +
theme_minimal()+
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
There does not seem to be a linear relationship between academic year length and PISA scores.

